---
title: "Final exam"
author: "Mohammad Hasan"
date: "February 8, 2023"
output:
  html_document: default
---
Question 2 (a) : Download the SPSS data file KiGGS03_06.sav from moodle and import it into R.
```{r}
library(foreign)

data <- read.spss("Data/KiGGS03_06.sav", to.data.frame=TRUE)
data

```
Question 2 (b) : Create a new dataframe in R named kiggs, which contains all variables (and only these) for
the analysis (E070M, E070V, E072, E074, arztZ01, sex, age2, schichtz, e065z)
```{r}
kiggs <- data[,c('E070M','E070V','E072','E074','arztZ01','sex','age2','schichtz','e065z')]
kiggs
```
Question 2 (c) : Run the formatting steps in the provided Rmd file data_formatting.Rmd. Save this
formatted dataframe on your computer, e.g. on your desktop.
```{r}
# Formatting the data
kiggs$E070M    <- factor(kiggs$E070M,    labels = c("yes, daily", "yes, occasionally", "no"))
kiggs$E070V    <- factor(kiggs$E070V,    labels = c("yes, daily", "yes, occasionally", "no"))
kiggs$E072     <- factor(kiggs$E072,     labels = c("yes, regularly", "yes, from time to time", "no, never"))
kiggs$E074     <- factor(kiggs$E074,     labels = c("yes, regularly", "yes, from time to time", "no, never", "has not breastfed"))
kiggs$sex      <- factor(kiggs$sex,      labels = c("boys", "girls"))
kiggs$age2     <- factor(kiggs$age2,     labels = c("0-1y", "2-3y", "4-5y", "6-7y", "8-9y", "10-11y", "12-13y", "14-15y", "16-17y"))
kiggs$schichtz <- factor(kiggs$schichtz, labels = c("low social status", "medium social status", "high social status"))

#Saving the data in my computer
save(data, file = "Data/kiggs.sav")

```

Question 3 (a) : Check that the variables E070M, E070V, E072, E074 are all factors. If they are not,
transform them into factors
```{r}
#Checking whether the variables E070M, E070V, E072, E074 are all factors or not
is.factor(kiggs$E070M)
is.factor(kiggs$E070V)
is.factor(kiggs$E072)
is.factor(kiggs$E074)
```
Question 3 (a) : Set the value "has not breastfed" of variable E074 to NA for all children.
```{r}
#Setting the value "has not breastfed" of variable E074 to NA for all children

kiggs["E074"][kiggs["E074"] == "has not breastfed"] <- "NA"

```
Question 3 (a) : Delete this now empty factor level from the variable
```{r}
kiggs
```






