```{r}
library("tidyverse")
#Loading the Data

students <- readxl::read_excel("Data/data_students.xlsx")

faculty <- readxl::read_excel("Data/data_faculty.xlsx")

```

```{r}
#Checking the data tables
students
faculty
```

```{r}
#Counting the NA values in Dataset
yes_count <- sum(students$age == "NA")
yes_count
```
```{r}
#Removing the rows with NA values
students_new<-students[!(students$age=="NA"),]
students_new
```
```{r}
#Checking agian whether NA still available or not
yes_count <- sum(students_new$age == "NA")
yes_count
```

So, we do not have any NA values in our database.

```{r}
# Counting the number of unique values in each column
unique_counts_students <- sapply(students_new, function(x) length(unique(x)))
unique_counts_faculty <- sapply(faculty, function(x) length(unique(x)))

# Viewing the counts
unique_counts_students
unique_counts_faculty
```
```{r}
#Checking the datatypes of the columns
str(students_new)
str(faculty)
```

The Data are in character form, So at first we have to make it numeric.

```{r}
# Converting the character columns to numeric

students_new$age <- as.numeric(students_new$age)

# Converting the country column to a factor
students_new$cob <- factor(students_new$cob)
#levels to see which number is assigned to which value
levels(students_new$cob)
# Converting the factor to a numeric vector
students_new$cob <- as.numeric(students_new$cob)


students_new$course <- factor(students_new$course)
students_new$course <- as.numeric(students_new$course)


students_new$faculty <- factor(students_new$faculty)
levels(students_new$faculty)
students_new$faculty <- as.numeric(students_new$faculty)


students_new$gpa_2010 <- as.numeric(students_new$gpa_2010)
students_new$gpa_2011 <- as.numeric(students_new$gpa_2011)
students_new$gpa_2012 <- as.numeric(students_new$gpa_2012)
students_new$gpa_2013 <- as.numeric(students_new$gpa_2013)
students_new$gpa_2014 <- as.numeric(students_new$gpa_2014)
students_new$gpa_2015 <- as.numeric(students_new$gpa_2015)
students_new$gpa_2016 <- as.numeric(students_new$gpa_2016)
students_new$gpa_2017 <- as.numeric(students_new$gpa_2017)
students_new$gpa_2018 <- as.numeric(students_new$gpa_2018)
students_new$gpa_2019 <- as.numeric(students_new$gpa_2019)
students_new$gpa_2020 <- as.numeric(students_new$gpa_2020)

students_new$job <- factor(students_new$job)
students_new$job <- as.numeric(students_new$job)

students_new$lifesat <- as.numeric(students_new$lifesat)
students_new$like <- as.numeric(students_new$like)

students_new$relationship <- factor(students_new$relationship)
levels(students_new$relationship)
students_new$relationship <- as.numeric(students_new$relationship)

students_new$sex <- factor(students_new$sex)
levels(students_new$sex)
students_new$sex <- as.numeric(students_new$sex)


students_new$term <- as.numeric(students_new$term)

students_new$university <- factor(students_new$university)
students_new$university <- as.numeric(students_new$university)


faculty$faculty <- factor(faculty$faculty)
levels(faculty$faculty)
faculty$faculty <- as.numeric(faculty$faculty)

# Checking the data types of the columns
str(students_new)

```




1.	Create a summary table with a) the percentage of non-German students, b) average life satisfaction, c) percentage of students in a relationship, d) sex ratio, e) percentage of students over 30, f) average gpa in 2010 and g) average number of terms. All variables should be displayed by faculty. 
```{r}
# Calculating the summary statistics by faculty
summary_table <- students_new %>%
  group_by(faculty) %>%
  summarise(
    Non_German = sum(cob != 3) / n() * 100, # 3 represents Germany
    Avg_Life_Satisfaction = mean(lifesat),
    In_Relationship = sum(relationship == 1) / n() * 100, #1 represents In a Relationship
    Sex_Ratio = sum(sex == 1) / sum(sex == 2),
    Over_30 = sum(age > 30) / n() * 100,
    Avg_GPA_2010 = mean(gpa_2010),
    Avg_Terms = mean(term)
  )

# Viewing the summary table
summary_table

```



2.	Create a graph showing differences in average life satisfaction by faculty and relationship status. 
```{r}
library(ggplot2)

# Calculating the average life satisfaction by faculty and relationship status
data_summary <- students_new %>%
  group_by(faculty, relationship) %>%
  summarise(Avg_Life_Satisfaction = mean(lifesat))

# Creating a bar chart showing the differences in average life satisfaction
ggplot(data_summary, aes(x = faculty, y = Avg_Life_Satisfaction, fill = relationship)) +
  geom_col(position = "dodge") +
  labs(x = "Faculty", y = "Average Life Satisfaction", fill = "Relationship Status")

```



3.	Combine faculty data with student data into one dataset. Show differences in the average cost of the career by faculty and job status.
```{r}
# Merging the student and faculty data by faculty
data <- merge(students_new, faculty, by = "faculty")

# Calculating the average cost by faculty and job status
data_summary <- data %>%
  group_by(faculty, job) %>%
  summarise(Avg_Cost = mean(costs))

# Creating a bar chart showing the differences in average cost
ggplot(data_summary, aes(x = faculty, y = Avg_Cost, fill = job)) +
  geom_col(position = "dodge") +
  labs(x = "Faculty", y = "Average Cost", fill = "Job Status")

```


4.	Create a graph showing the association between life satisfaction and age.
```{r}
# Creating a scatter plot showing the association between life satisfaction and age
ggplot(students_new, aes(x = age, y = lifesat)) +
  geom_point() +
  labs(x = "Age", y = "Life Satisfaction")

```


5.	Create a dummy variable for students with a gpa above 2 for each year (using a loop). 
```{r}
# Creating a dummy variable for each year
for (year in 2010:2020) {
  # Creating the column name for the current year
  column_name <- paste0("gpa_", year)
  
  # Creating the dummy variable for the current year
  students_new[[paste0(column_name, "_above_2")]] <- ifelse(students_new[[column_name]] > 2, 1, 0)
}

# Viewing the first few rows of the data
head(students_new)

```


Linear regression
1.	Apply Linear Regression models to test whether there is a statistically significant relationship between life satisfaction and relationship status (controlling for age and sex).
```{r}
# Fitting a linear regression model
model <- lm(lifesat ~ relationship + age + sex, data = students_new)

# Viewing the summary of the model
summary(model)

```
In the summary of the model, the p-value associated with the relationship variable is 0.2945, which is greater than the common significance level of 0.05. This means that we cannot reject the null hypothesis that there is no relationship between life satisfaction and relationship status while controlling for age and sex.





2.	Run appropriate regression diagnostics tests

```{r}
# Plotting the residuals vs. fitted values to check for linearity and homoscedasticity
plot(model, which = 1)

# Plotting the QQ plot of residuals to check for normality
plot(model, which = 2)

# Plotting the scale-location plot to check for homoscedasticity
plot(model, which = 3)

# Plotting the residuals vs. leverage plot to check for influential observations
plot(model, which = 5)

```




3.	Plot the effects of relationship status, age and gender using a coefficient plot
```{r}
# loading the arm package
library(arm)


# Plotting the coefficients using coefplot
coefplot(model)

```


4.	Provide a formatted regression table for the overall model output
```{r}
#loading the stargazer package
library(stargazer)

# Creating a formatted regression table using stargazer
stargazer(model, type = "text")

```


5.	Predict the average life satisfaction for male students in a relationship and, separately, the average life satisfaction for female students. 
```{r}
# Using the predict function to make predictions
predictions <- predict(model)

# Adding the predictions to the students_new dataset
students_new$predicted_lifesat <- predictions

# Calculating the mean predicted life satisfaction for each group
aggregate(predicted_lifesat ~ relationship + sex, data = students_new, FUN = mean)



```
Here, 1 Represent Female and 2 Represents Male
And, 1 Represent In a Relationship and 2 Represents Single





Logistic regression

1.	Columns I to O represent information on Grade points averages (GPA) by year. Dichotomize these variables (scores below 2 are “good”, scores below 2 are “bad”).
```{r}
# Create new binary variables for each GPA column
students_new$gpa_2010_bin <- ifelse(students_new$gpa_2010 < 2, "good", "bad")
students_new$gpa_2011_bin <- ifelse(students_new$gpa_2011 < 2, "good", "bad")
students_new$gpa_2012_bin <- ifelse(students_new$gpa_2012 < 2, "good", "bad")
students_new$gpa_2013_bin <- ifelse(students_new$gpa_2013 < 2, "good", "bad")
students_new$gpa_2014_bin <- ifelse(students_new$gpa_2014 < 2, "good", "bad")
students_new$gpa_2015_bin <- ifelse(students_new$gpa_2015 < 2, "good", "bad")
students_new$gpa_2016_bin <- ifelse(students_new$gpa_2016 < 2, "good", "bad")
students_new$gpa_2017_bin <- ifelse(students_new$gpa_2017 < 2, "good", "bad")
students_new$gpa_2018_bin <- ifelse(students_new$gpa_2018 < 2, "good", "bad")
students_new$gpa_2019_bin <- ifelse(students_new$gpa_2019 < 2, "good", "bad")
students_new$gpa_2020_bin <- ifelse(students_new$gpa_2020 < 2, "good", "bad")

```


2.	Use logistic regression to test the association between having a “good” GPA and faculty, sex and age.

```{r}
# Creating a new binary variable for gpa_2010 where value is 1 if the student’s GPA in 2010 was below 2 and a value of 0 otherwise. 
students_new$gpa_2010_bin <- ifelse(students_new$gpa_2010 < 2, 1, 0)

# Fitting a logistic regression model
model <- glm(gpa_2010_bin ~ faculty + sex + age, data = students_new, family = binomial)

# Viewing the summary of the model
summary(model)


```

3.	Estimate the model for all available years and report the results in one graph.

```{r}
library(broom)
library(ggplot2)
library(arm)
# Create new binary variables for each GPA column
students_new$gpa_2010_bin <- ifelse(students_new$gpa_2010 < 2, 1, 0)
students_new$gpa_2011_bin <- ifelse(students_new$gpa_2011 < 2, 1, 0)
students_new$gpa_2012_bin <- ifelse(students_new$gpa_2012 < 2, 1, 0)
students_new$gpa_2013_bin <- ifelse(students_new$gpa_2013 < 2, 1, 0)
students_new$gpa_2014_bin <- ifelse(students_new$gpa_2014 < 2, 1, 0)
students_new$gpa_2015_bin <- ifelse(students_new$gpa_2015 < 2, 1, 0)
students_new$gpa_2016_bin <- ifelse(students_new$gpa_2016 < 2, 1, 0)
students_new$gpa_2017_bin <- ifelse(students_new$gpa_2017 < 2, 1, 0)
students_new$gpa_2018_bin <- ifelse(students_new$gpa_2018 < 2, 1, 0)
students_new$gpa_2019_bin <- ifelse(students_new$gpa_2019 < 2, 1, 0)
students_new$gpa_2020_bin <- ifelse(students_new$gpa_2020 < 2, 1, 0)

# Creating a vector of the binary GPA variables
gpa_vars <- c("gpa_2010_bin", "gpa_2011_bin", "gpa_2012_bin", "gpa_2013_bin", "gpa_2014_bin", "gpa_2015_bin", "gpa_2016_bin", "gpa_2017_bin", "gpa_2018_bin", "gpa_2019_bin", "gpa_2020_bin")

# Fitting a separate logistic regression model for each year
models <- lapply(gpa_vars, function(x) {
    glm(as.formula(paste0(x," ~ faculty + sex + age")), data = students_new,family = binomial)
})

# Extracting the estimated coefficients from each model
coefs <- sapply(models, function(x) coef(x))
rownames(coefs) <- c("(Intercept)", "faculty", "sexMale", "age")

# Creating a line plot that shows the estimated coefficients for each year
matplot(t(coefs), type = "l", lty = 1, col = 1:4, ylab = "Estimated Coefficient")
legend("topleft", legend = rownames(coefs), col = 1:4, lty = 1)


```

